---
title: "Exercise 3.2 Missing Values"
author: "Sean Baron"
date: "3/27/2018"
output: html_document
---

```{r results = FALSE, echo = FALSE, message = FALSE}
install.packages("gdata", repos = "http://cran.us.r-project.org")
install.packages("dplyr", repos = "http://cran.us.r-project.org")
library("gdata")
library("dplyr")
```


### Exercise Description

Go through various exercises and methods for dealing with missing values.

### Step 0: Load data into R.

```{r}
df <- read.xls("titanic3.xls", sheet = 1, header = TRUE)
```

### Step 1: Port of Embarkation 

Change emtpy values in embarked to the factor level S:

```{r}
df$embarked[which(df$embarked == "")] <- as.factor("S")
```

### Step 2: Age

Replace all missing age values (NAs) with mean of age column.

```{r}
df$age[which(is.na(df$age))] <- mean(df$age, na.rm = TRUE)
```

Alternatively, we could've used a linear regression to approximate each of the missing values.

### Step 3: Lifeboat

Change emtpy values for 'r boat' to NA.

```{r}
df$boat[which(df$life == "")] <- NA
```

### Step 4: Cabin

There are a number of missing cabin numbers. The only case where it would make sense to fill these missing values is if we had a accurate model for classifying what those cabin numbers would be.

A significant number of cabin values are missing (`r round(100*(sum(df$cabin == "")/length(df$cabin)), digits = 2)`%). Given the significant number of missing values and the possiblility that the majority are missing, it's worth seeing if missing or no missing is a predictor of survival...

Create a missing or not missing dummy variable for cabin.

Change empty values for 'r boat' to NA.
```{r}
df$has_cabin_number <- ifelse(!df$cabin == "", 1, 0)
```

